---
title: "Presentation"
author: "Group 14"
date: "April 24, 2017"
output: html_document
---

```{r setup, include=FALSE}
library(jsonlite)
library(stringr)
library(tidyverse)
library(readr)


data <- read_csv("14_data.csv")
```


####Plot 1  
Slightly skewed indicating that younger people may be more likely to be killed by police.     
```{r}
data_age <- na.omit(select(data,age))
ggplot(data_age) + geom_histogram(aes(x=age), fill = "#FF9999", color = "black", binwidth = 4) + 
    ggtitle('Police Killings by Age') + labs(x = 'Age', y = '# of people killed by police') +
    theme(panel.background = element_blank(), 
          axis.line = element_line(colour = "black", size=.1))+
    scale_x_continuous(breaks=c(0,20, 40, 60, 80),
                       limits=c(0,90),
                       expand=c(0, 1)) +
    scale_y_continuous(breaks=c(0, 50, 100, 150),
                       limits=c(0, 150),
                       expand=c(0,0)) 
```

####Plot 2    
```{r}
race_count <- data %>% group_by(raceethnicity) %>% count()
ggplot(race_count) + geom_bar(aes(x=raceethnicity, y=n), stat = "identity", fill="#FF9999") + 
    theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1), axis.line = element_line(colour = "black", size=.1),
          panel.background = element_blank()) +
    ggtitle('Police Killings by Race/Ethnicity') +
    labs(x = 'Race/Ethnicity', y = '# of people killed by police')
```

####Plot 3  
The distribution among race appears to be different for those who are unarmed than for all people.  
```{r }
race_count <- data %>% filter(is_armed==0) %>% group_by(raceethnicity) %>% count()
ggplot(race_count) + geom_bar(aes(x=raceethnicity, y=n), stat = "identity", fill="#FF9999") + 
    theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1), axis.line = element_line(colour = "black", size=.1),
          panel.background = element_blank()) +
    ggtitle('Police Killings of Unarmed People') +
    labs(x = 'Race/Ethnicity', y = '# of people killed by police')
```






####Plot 4  
There appears to be a cycle over the year.  

```{r}
date_count <- data %>% group_by(month_num) %>% count()
ggplot(date_count) + 
    geom_bar(aes(x=month_num, y=n), stat = "identity", fill="#FF9999") + 
    ggtitle('Police Killings in 2015') +
    labs(x = 'Month', y = '# of people killed by police') + 
    theme(axis.line = element_line(colour = "black", size=.1), 
          panel.background = element_blank()) +
    scale_x_discrete(labels=c("Jan","Feb","Mar","Apr","May","Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
```

####Plot 5   
```{r ,warning=FALSE}
avgGrad <- data %>% group_by(state) %>% summarise(avgPercentHighSchoolGrad = mean(as.numeric(percent_25_or_older_HighSchool_grads), na.rm = TRUE))
stateCount <- data %>% group_by(state) %>% summarise(DeathsPerState = n())
gradsCount <- left_join(stateCount, avgGrad, by = c('state' = 'state'))


ggplot(gradsCount) + 
    geom_point(mapping = aes(x = avgPercentHighSchoolGrad, y = DeathsPerState), color="#FF9999") +
    ggtitle('Death counts per state are higher in states with lower percentage graduation rates') +
    theme(axis.line = element_line(colour = "black", size=.1), 
          panel.background = element_blank()) +
    labs(x = 'State average graduation rate', y = '# of people killed by police per state') 
```

####Plot 6 
Males are killed by police much more often than females. 
```{r}
monthCount <- data %>% group_by(month, gender) %>% summarise(DeathsPerMonth = n())

# plot the total number of deaths per month, and 
ggplot(data) + geom_bar(mapping = aes(x=month, fill = gender )) + 
    scale_fill_manual(values = c('hotpink', 'blue', 'green')) +
    theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
    ggtitle('Number of People Killed By Police Officers in 2015 by Month')
```


####Plot 7  
Most police killings occurred in the South and the West.   
```{r}
region_count <- data %>% group_by(region) %>% count()
ggplot(region_count) + 
    geom_bar(aes(x=region, y=n), stat = "identity", fill="#FF9999") + 
    ggtitle('Police Killings by Region') +
    labs(x = 'Region', y = '# of people killed by police') + 
    theme(axis.line = element_line(colour = "black", size=.1), 
          panel.background = element_blank())
```


####Plot 8  
```{r, warning=FALSE}
Death_Type <- factor(data$classification)
ggplot(data) + geom_boxplot(aes(x=classification, y=age, fill = Death_Type)) + ggtitle('Boxplot of Age per Death Type') +
labs(x = 'Death Type', y = 'Age') + 
scale_x_discrete(limit = c("Death in custody", "Gunshot", "Struck by vehicle", "Taser"))
```


***

##Stats

```{r}
#Average age of those killed by police
mean_age <- mean(na.omit(data$age))
mean_age

median_age <- median(na.omit(data$age))
median_age

#Average age of those killed by police - Race/Ecthnicity = White
w_data <- data %>% filter(raceethnicity=="White")   
mean_age_w <- mean(na.omit(w_data$age))
mean_age_w

#Average age of those killed by police - Race/Ecthnicity = Black
b_data <- data %>% filter(raceethnicity=="Black")   
mean_age_b <- mean(na.omit(b_data$age))
mean_age_b

#Average age of those killed by police - Race/Ecthnicity = Hispanic/Latino
h_data <- data %>% filter(raceethnicity=="Hispanic/Latino")   
mean_age_h <- mean(na.omit(h_data$age))
mean_age_h

#The mean age of those killed by police is younger for Race/Ethnicity = Black than White, and lowest for Hispanic/Latino



#Data summary
data_summary <- summary(data)

```





