---
title: "Project"
author: "Group 14"
date: "April 24, 2017"
output: html_document
---


```{r setup, warning=FALSE, message=F}
library(jsonlite)
library(stringr)
library(tidyverse)
library(readr)
data <- read_csv("guardian_data.csv")
# guardian_data is missing part of December, but has lat/long information. We found a second data (the-counted-2015.csv) set with all of December, but is missing lat/long. We will use lat_long_finder.Rmd to get latitude and longitude for December of the countede data set, and then join December to the table contained in guardian data.
data <- data %>%filter(month!="December")
decData <- read_csv('decDataComplete.csv')
data <- rbind(data,decData)
```


Get tract 
```{r}
row_to_url <- function(row){
    # set the URL base
    basestring = "https://www.broadbandmap.gov/broadbandmap/census/tract?"
    endstring = "&format=json"
    
    x = paste0("latitude=", row$lat)
    y = paste0("&", "longitude=", row$long)
    
    # create url string from base, coordinate, and end parts        
    urlstring = paste0(basestring, x, y, endstring)
    
    return(urlstring)
}

row_to_url(data[5, ])
```


Now lets request some sample data. 
```{r, cache=F}
url_1 = row_to_url(data[1, ])
url_1

#We can use the `readLines( )` function to retrieve data from the API by calling `readLines(aURL)`.
dat = readLines(url_1)
dat

#The value `dat` is in string format, we use the `jsonlite` to package to convert.
datdf = fromJSON(dat)
datdf

#The Results can be accessed as follows.
datdf$Results$censusTract

#Fips
datdf$Results$censusTract['fips']

#And there appears to be a useful location type field here.
#datdf$results$geometry$location_type
```



Functions to run all at once
```{r, warning=FALSE}
get_and_parse <- function(urlstring) {
    dat = readLines(urlstring)
    datdf = fromJSON(dat)
    return(datdf)
}

add_row <- function(df, jsndat, reqdat) {
    # expects df to have columns [street, city, state, lat, lon, loctype]
    # req dat is a row of data used for request [street, city, state]
    resdf = data.frame(
        name = reqdat$name,
        age =  reqdat$age,
        gender = reqdat$gender,               
        raceethnicity = reqdat$raceethnicity,       
        month = reqdat$month,              
        day = reqdat$day,
        year = reqdat$year,                
        streetaddress = reqdat$streetaddress,      
        city = reqdat$city,                
        state = reqdat$state,              
        classification = reqdat$classification,      
        lawenforcementagency = reqdat$lawenforcementagency,
        lat = reqdat$lat,                 
        long = reqdat$long,                
        armed = reqdat$armed, 
        state_fp = jsndat$Results$censusTract$stateFips,
        geo_id = jsndat$Results$censusTract$fips,
        census_tract = jsndat$Results$censusTract$name
    )
    return(rbind(df, resdf))
}

map_data <- function(df) {
    # expects df to have columns [street, city, state]
    # initialize a dataframe with appropriate columns using the first row of
    # our request data
    reqrow = df[1,]

    qryurl = row_to_url(reqrow)
    jsndf = get_and_parse(qryurl)
    
    newdf = data.frame(
        name = reqrow$name,
        age =  reqrow$age,
        gender = reqrow$gender,               
        raceethnicity = reqrow$raceethnicity,       
        month = reqrow$month,              
        day = reqrow$day,
        year = reqrow$year,                
        streetaddress = reqrow$streetaddress,      
        city = reqrow$city,                
        state = reqrow$state,              
        classification = reqrow$classification,      
        lawenforcementagency = reqrow$lawenforcementagency,
        lat = reqrow$lat,                 
        long = reqrow$long,                
        armed = reqrow$armed, 
        state_fp = jsndf$Results$censusTract$stateFips,
        geo_id = jsndf$Results$censusTract$fips,
        census_tract = jsndf$Results$censusTract$name
    )

    # populate the dataframe with the rest of the request data
    for (ix in 2:nrow(df)) {
        tryCatch({
            reqrow = df[ix, c("name","age","gender","raceethnicity","month","day","year","streetaddress","city","state", "classification","lawenforcementagency","lat","long","armed" )]

            if (!is.null(reqrow[13]) | !is.null(reqrow[14])){
                qryurl <-  row_to_url(reqrow)
                jsndf <-  get_and_parse(qryurl)
                newdf <-  add_row(newdf, jsndf, reqrow)
            }
        })
    }

    return(newdf)
}

#shortDatacomplete <- map_data(shortData)
completeData <- map_data(data)
```

Adds a column to dataframe giving the county of incidentusing googles geocoding API
```{r}
countyFinder <- function(row){
    
    address <- row[1]
    city <- row[2]
    state <- row[3]
    googleKey <-'&key=AIzaSyDZ0xPO1ebMXg1p7KqvBtsUGp2lIbBJLTQ' ## User needs to add their own Google API Key here as string
    url <- paste0("https://maps.googleapis.com/maps/api/geocode/json?address=", address, ",+",
                  city, ",+", state, googleKey)
    
    raw_json <- readLines(url)
    
    # turn the json into a nice list
    info <- fromJSON(raw_json)
    
    # Get the first element of from the results list, and then the first componenent of address component
    components <- info["results"][[1]][["address_components"]][[1]]
    # Check component is not null
    if(!is.null(components)){
        # Check each elelement in address component to find the one that is listed as "administrative_area_level_2"
        # This component will give us the county name for a given request
        for (i in 1:nrow(components)){
            if (str_detect(components[i,]["types"], "administrative_area_level_2")){
                return(unlist(components[i,][["long_name"]]))
            }
        }
    }
}

importantCols <- dplyr::select(data, streetaddress, city, state)
counties <- apply(importantCols, 1, countyFinder)
counties1 <- as.character(counties)
completeData <- add_column(completeData,county = (counties1))
write_csv(completeData, 'data_missing_counties.csv')
```
Note: Some counties will appear ass null, preventing us from getting further information. From this point, we will hand code in the remaining counties that were left out. The file below has the hand entered data
```{r}
completeData <- read_csv('all_months_with_counties.csv')
completeData <- mutate(completeData, census_tract = as.character(census_tract))
completeData <- mutate(completeData, state_fp = as.character(state_fp))
```

Now that we have tracts, we get demographic and finacial information using the census API
```{r, warning=FALSE}
#Initialize data frames for important variables
medianIncomeByTract <- data.frame()


# Create base url string
    start_url <- 'http://api.census.gov/data/2015/acs5/profile?get='
    end_url <- ',NAME&for=tract:*&in=state:'
    key <- '&key=913ff64a27817f08554c0606053c0715f1de4a06'

     
# Get median Household Income    
for (i in 1:56){
      tryCatch({
        medianIncomeJSON <- readLines(paste0(start_url,'DP03_0062E',end_url,i,key))
        medianIncome <- fromJSON(medianIncomeJSON)
        medianIncome <- data.frame(medianIncome)
        medianIncome <- filter(medianIncome, X2 != 'NAME')
        medianIncome <- mutate(medianIncome, tract = unlist(lapply(str_extract_all(X2, "[0-9\\.]+"), `[[`, 1))) 
        medianIncome <- mutate(medianIncome, county1 = unlist(lapply(str_extract_all(X2, "(?<=,)[^,]+(?=,)"), `[[`, 1))) %>% select(X1, tract, county1, X3) 
        medianIncomeByTract <- rbind.data.frame(medianIncomeByTract, medianIncome)
    }, error = function(err) {
        i <- i+1
    })
}
medianIncomeByTractTest <-  mutate(medianIncomeByTract,county = str_trim(county1), X3 = as.character(X3))
medianIncomeByTractTest <-  mutate(medianIncomeByTractTest, X3 = substr(X3,regexpr("[^0]",X3),nchar(X3)))

completeData <- left_join(completeData, medianIncomeByTractTest, by=c('census_tract'='tract', 'state_fp'= 'X3', "county" = "county")) %>% select(-county1)
names(completeData)[names(completeData) == 'X1'] <- 'median_household_Income'


```

Now we find the demographics of each census tract by looking at the breakdown of race percentages. Note, we are not using all races listed on the census, and we are also double counting the people that identify as more than one race
```{r, warning=FALSE}
# Get percent white population  
percentWhiteByTract <- data.frame()
for (i in 1:56){
      tryCatch({
        percentWhiteJSON <- readLines(paste0(start_url,'DP05_0072PE',end_url,i,key))
        percentWhite <- fromJSON(percentWhiteJSON)
        percentWhite <- data.frame(percentWhite)
        percentWhite <- filter(percentWhite, X2 != 'NAME')
        percentWhite <- mutate(percentWhite, tract = unlist(lapply(str_extract_all(X2, "[0-9\\.]+"), `[[`, 1))) 
        percentWhite <- mutate(percentWhite, county1 = unlist(lapply(str_extract_all(X2, "(?<=,)[^,]+(?=,)"), `[[`, 1))) %>% select(X1, tract, county1, X3) 
        percentWhiteByTract <- rbind.data.frame(percentWhiteByTract,percentWhite)
    }, error = function(err) {
        i <- i+1
    })
}

percentWhiteByTractTest <-  mutate(percentWhiteByTract,county = str_trim(county1), X3 = as.character(X3))
percentWhiteByTractTest <-  mutate(percentWhiteByTractTest, X3 = substr(X3,regexpr("[^0]",X3),nchar(X3)))

completeData <- left_join(completeData, percentWhiteByTractTest, by=c('census_tract'='tract', 'state_fp'= 'X3', "county" = "county")) %>% select(-county1)
names(completeData)[names(completeData) == 'X1'] <- 'percent_white_population'
```


```{r, warning=FALSE}
# Get percent black population  

percentBlackByTract <- data.frame()
for (i in 1:56){
      tryCatch({
        percentBlackJSON <- readLines(paste0(start_url,'DP05_0033PE',end_url,i,key))
        percentBlack <- fromJSON(percentBlackJSON)
        percentBlack <- data.frame(percentBlack)
        percentBlack <- filter(percentBlack, X2 != 'NAME')
        percentBlack <- mutate(percentBlack, tract = unlist(lapply(str_extract_all(X2, "[0-9\\.]+"), `[[`, 1))) 
        percentBlack <- mutate(percentBlack, county1 = unlist(lapply(str_extract_all(X2, "(?<=,)[^,]+(?=,)"), `[[`, 1))) %>% select(X1, tract, county1, X3) 
        percentBlackByTract <- rbind.data.frame(percentBlackByTract,percentBlack)
    }, error = function(err) {
        i <- i+1
    })
}
percentBlackByTractTest <-  mutate(percentBlackByTract,county = str_trim(county1), X3 = as.character(X3))
percentBlackByTractTest <-  mutate(percentBlackByTractTest, X3 = substr(X3,regexpr("[^0]",X3),nchar(X3)))

completeData <- left_join(completeData, percentBlackByTractTest, by=c('census_tract'='tract', 'state_fp'= 'X3', "county" = "county")) %>% select(-county1)
names(completeData)[names(completeData) == 'X1'] <- 'percent_black_population'
```


```{r, warning=FALSE}
# Get percent Hispanic population  


percentHispanicByTract <- data.frame()

for (i in 1:56){
      tryCatch({
        percentHispanicJSON <- readLines(paste0(start_url,'DP05_0066PE',end_url,i,key))
        percentHispanic <- fromJSON(percentHispanicJSON)
        percentHispanic <- data.frame(percentHispanic)
        percentHispanic <- filter(percentHispanic, X2 != 'NAME')
        percentHispanic <- mutate(percentHispanic, tract = unlist(lapply(str_extract_all(X2, "[0-9\\.]+"), `[[`, 1))) 
        percentHispanic <- mutate(percentHispanic, county1 = unlist(lapply(str_extract_all(X2, "(?<=,)[^,]+(?=,)"), `[[`, 1))) %>% select(X1, tract, county1, X3) 
        percentHispanicByTract <- rbind.data.frame(percentHispanicByTract,percentHispanic)
    }, error = function(err) {
        i <- i+1
    })
}
percentHispanicByTractTest <-  mutate(percentHispanicByTract,county = str_trim(county1), X3 = as.character(X3))
percentHispanicByTractTest <-  mutate(percentHispanicByTractTest, X3 = substr(X3,regexpr("[^0]",X3),nchar(X3)))

completeData <- left_join(completeData, percentHispanicByTractTest, by=c('census_tract'='tract', 'state_fp'= 'X3', "county" = "county")) %>% select(-county1)
names(completeData)[names(completeData) == 'X1'] <- 'percent_hispanic_population'
```

```{r, warning = FALSE}
# Get percent of population that is Native American
percentNativeByTract <- data.frame()

for (i in 1:56){
      tryCatch({
        percentNativeJSON <- readLines(paste0(start_url,'DP05_0074PE',end_url,i,key))
        percentNative <- fromJSON(percentNativeJSON)
        percentNative <- data.frame(percentNative)
        percentNative <- filter(percentNative, X2 != 'NAME')
        percentNative <- mutate(percentNative, tract = unlist(lapply(str_extract_all(X2, "[0-9\\.]+"), `[[`, 1))) 
        percentNative <- mutate(percentNative, county1 = unlist(lapply(str_extract_all(X2, "(?<=,)[^,]+(?=,)"), `[[`, 1))) %>% select(X1, tract, county1, X3) 
        percentNativeByTract <- rbind.data.frame(percentNativeByTract,percentNative)
    }, error = function(err) {
        i <- i+1
    })
}

percentNativeByTractTest <-  mutate(percentNativeByTract,county = str_trim(county1), X3 = as.character(X3))
percentNativeByTractTest <-  mutate(percentNativeByTractTest, X3 = substr(X3,regexpr("[^0]",X3),nchar(X3)))

completeData <- left_join(completeData, percentNativeByTractTest, by=c('census_tract'='tract', 'state_fp'= 'X3', "county" = "county")) %>% select(-county1)
names(completeData)[names(completeData) == 'X1'] <- 'percent_native_population'

```


```{r, warning=FALSE}
percentHSGradsByTract <- data.frame()
# Get percent of people aged 25 or older that are high school graduate or higher
for (i in 1:56){
      tryCatch({
        percentHSGradsJSON <- readLines(paste0(start_url,'DP02_0066PE',end_url,i,key))
        percentHSGrads <- fromJSON(percentHSGradsJSON)
        percentHSGrads <- data.frame(percentHSGrads)
        percentHSGrads <- filter(percentHSGrads, X2 != 'NAME')
        percentHSGrads <- mutate(percentHSGrads, tract = unlist(lapply(str_extract_all(X2, "[0-9\\.]+"), `[[`, 1))) 
        percentHSGrads <- mutate(percentHSGrads, county1 = unlist(lapply(str_extract_all(X2, "(?<=,)[^,]+(?=,)"), `[[`, 1))) %>% select(X1, tract, county1, X3) 
        percentHSGradsByTract <- rbind.data.frame(percentHSGradsByTract,percentHSGrads)
    }, error = function(err) {
        i <- i+1
    })
}
percentHSGradsByTractTest <-  mutate(percentHSGradsByTract,county = str_trim(county1), X3 = as.character(X3))
percentHSGradsByTractTest <-  mutate(percentHSGradsByTractTest, X3 = substr(X3,regexpr("[^0]",X3),nchar(X3)))

completeData <- left_join(completeData, percentHSGradsByTractTest, by=c('census_tract'='tract', 'state_fp'= 'X3', "county" = "county")) %>% select(-county1)
names(completeData)[names(completeData) == 'X1'] <- 'percent_25_or_older_HighSchool_grads'
```


Write finished data set to csv
```{r}
write_csv(completeData, '14_data.csv')
#completeData <- read_csv('14_data.csv')
```


