---
title: "process_notebook"
author: "Group 14"
date: "5/4/2017"
output: html_document
---
```{r setup, warning=FALSE, message=F, echo=FALSE}
library(jsonlite)
library(stringr)
library(tidyverse)
library(readr)

data <- read_csv('14_data_3.csv')
```


# Abstract: 
One paragraph at the very beginning of the document summarizing everything.
```{r}

```


# Overview and Motivation: 
Provide an overview of the project goals and the motivation for it. Consider that this will be read by people who did not see your project proposal.
```{r}

```


# Related Work: 
Anything that inspired you, such as a paper, a web site, or something we discussed in class.
```{r}

```


# Initial Questions: 
What questions are you trying to answer? 
How did these questions evolve over the course of the project? 
What new questions did you consider in the course of your analysis? 


# Data: 
Source, scraping method, cleanup, storage, etc.
```{r}

```


# Exploratory Data Analysis: 
What visualizations did you use to look at your data in different ways? 
What are the different statistical methods you considered? 
Justify the decisions you made, and show any major changes to your ideas. 
How did you reach these conclusions?
```{r}

```


## Time 
The following plot shows the number of people killed by police during each month of 2015.  
There appears to be somewhat of a cycle in which the numer of deaths per month rises to a peak and then begins to decline before rising again. 
This could possibly be due to the nature of media coverage or the time of year. 
```{r, echo = FALSE}
date_count <- data %>% group_by(month_num) %>% count()
ggplot(date_count) + 
    geom_line(aes(x=month_num, y=n), color="#FF9999") +
    geom_point(aes(x=month_num, y=n), stat = "identity") + 
    ggtitle('Police Killings in 2015') +
    labs(x = 'Month', y = '# of people killed by police') + 
    theme(axis.line.x = element_line(color = "black", size=.1),
        axis.line.y = element_line(color="black", size = .1), 
          panel.background = element_blank()) +
    scale_x_discrete(limits=c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12"),
                     labels=c("Jan","Feb","Mar","Apr","May","Jun", "Jul", "Aug", "Sep", "Oct", "Nov","Dec"))  
```

## Age

## Gender

## Race 
```{r, echo=FALSE}
pop_2015 = 321418820

asian_share = 0.056
black_share = 0.133
hispanic_share = 0.176
native_share = 0.012
white_share = 0.616

per_million_counts <- function(race_count){
    race_count$per_million <- NA
    for (i in 1:nrow(race_count)){
        if (race_count$raceethnicity[i] == "Asian/Pacific Islander"){
            race_count$per_million[i] = ((race_count$n[race_count$raceethnicity=="Asian/Pacific Islander"])*1000000)/(pop_2015*asian_share)
        } else if (race_count$raceethnicity[i] == "Black"){
            race_count$per_million[i] =     ((race_count$n[race_count$raceethnicity=="Black"])*1000000)/(pop_2015*black_share)
        } else if (race_count$raceethnicity[i] == "Hispanic/Latino"){ 
            race_count$per_million[i] = ((race_count$n[race_count$raceethnicity=="Hispanic/Latino"])*1000000)/(pop_2015*hispanic_share)
        } else if (race_count$raceethnicity[i] == "Native American"){ 
            race_count$per_million[i] = 
                ((race_count$n[race_count$raceethnicity=="Native American"])*1000000)/(pop_2015*native_share)
        } else if (race_count$raceethnicity[i] == "White"){ 
            race_count$per_million[i] = ((race_count$n[race_count$raceethnicity=="White"])*1000000)/(pop_2015*white_share)
        } else race_count$per_million[i] = NA       
    }
    return(race_count)
}
```

```{r, echo=FALSE}
race_count <- data %>% group_by(raceethnicity) %>% count()
race_count <- per_million_counts(race_count)
# Make a histogram
ggplot(na.omit(race_count)) + geom_bar(aes(x=raceethnicity, y=per_million), stat = "identity", fill="#FF9999") + 
    theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1), axis.line = element_line(colour = "black", size=.1),
          panel.background = element_blank()) +
    ggtitle('People Killed by Police by Race/Ethnicity') +
    labs(x = 'Race/Ethnicity', y = '# killed (per million)')
```

```{r, echo=FALSE}
race_count_armed <- data %>% filter(is_armed==0) %>% group_by(raceethnicity) %>% count()
race_count_armed <- per_million_counts(race_count_armed)
# Make a histogram
ggplot(na.omit(race_count_armed)) + geom_bar(aes(x=raceethnicity, y=per_million), stat = "identity", fill="#FF9999") + 
    theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1), axis.line = element_line(colour = "black", size=.1),
          panel.background = element_blank()) +
    ggtitle('Unarmed People Killed by Police by Race/Ethnicity') +
    labs(x = 'Race/Ethnicity', y = '# killed (per million)')
```





# Final Analysis: 
What did you learn about the data? 
How did you answer the questions? 
How can you justify your answers?
```{r}

```






