---
title: "Adding Variables"
author: "Group 14"
date: "5/4/2017"
output: html_document
---

```{r setup, warning=FALSE, message=F}
library(jsonlite)
library(stringr)
library(tidyverse)
library(readr)

data <- read_csv('guardian_data.csv')
completeData <- read_csv("completeData.csv")
```

# Adding Variables  
We added a number of variables to help with our analysis and visualization. 


## "Region" variable  
Divides the country into four regions.
```{r}
add_regions <- function(data){
    data$region <- NA
    
    NE = c("ME","NH","VT","MA","RI","CT","NY","PA","NJ", "DE")
    MW = c("IL", "IN", "MI", "OH", "WI", "IA", "KS", "MN", "MO", "NE", "ND", "SD")
    South = c("FL", "GA", "MD", "NC", "SC", "VA", "DC", "WV", "AL", "KY", "MS", "TN", "AR", "LA", "OK", "TX")
    West = c("AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY", "AK", "CA", "HI", "OR", "WA")
    
    data$region[data$state %in% NE] = "Northeast"
    data$region[data$state %in% MW] = "Midwest"
    data$region[data$state %in% South] = "South"
    data$region[data$state %in% West] = "West"    
    return(data)
}
```


## "Armed" Variable
Creates a binary variables about whether the victim was armed (1) or not (0). 
```{r, warning=FALSE}
add_armed <- function(data){
    data$is_armed <- ifelse(data$armed == "No", 0, 1)
    return(data)
}
```




## Racial majority variables
Creates variables indicating if the population of an area has a white, black, or hispanic majority.
```{r}
add_majority <- function(data){
    data$white_majority <- ifelse(data$percent_white_population >= 50, 1, 0)
    data$black_majority <- ifelse(data$percent_black_population >= 50, 1, 0)
    data$hispanic_majority <- ifelse(data$percent_hispanic_population >= 50, 1, 0)
    
    data <- mutate(data, in_majority = ifelse(data$raceethnicity == "White" 
                                              & data$white_majority == 1, "Yes",
                                              ifelse(data$raceethnicity == "Black" &
                                                         data$black_majority == 1, "Yes",
                                                     ifelse(data$raceethnicity == "Hispanic/Latino" 
                                                            & data$hispanic_majority == 1, "Yes",
                                                            ifelse(data$raceethnicity == "Unknown",
                                                                   NA,
                                                                   "No")))))
    return(data)
}
```


## "Date" variable
Creates a date variable that combines month, date, and year.
```{r}
make_date <- function(data){
    #format months
data$month_num <- NA
for (i in 1:nrow(data)){
if (data$month[i] == "January"){
    data$month_num[i] = "01"
}else if (data$month[i] == "February"){
    data$month_num[i] = "02"
    }else if (data$month[i] == "March"){
    data$month_num[i] = "03"
    }else if (data$month[i] == "April"){
    data$month_num[i] = "04"
    }else if (data$month[i] == "May"){
    data$month_num[i] = "05"
    }else if (data$month[i] == "June"){
    data$month_num[i] = "06"
    }else if (data$month[i] == "July"){
    data$month_num[i] = "07"
    }else if (data$month[i] == "August"){
    data$month_num[i] = "08"
    }else if (data$month[i] == "September"){
    data$month_num[i] = "09"
    }else if (data$month[i] == "October"){
    data$month_num[i] = "10"
    }else if (data$month[i] == "November"){
    data$month_num[i] = "11"
} else 
    data$month_num[i] = "12"
}


    #format days
data$day_num <- NA
for (i in 1:nrow(data)){
if (data$day[i] == "1"){
    data$day_num[i] = "01"
}else if (data$day[i] == "2"){
    data$day_num[i] = "02"
    }else if (data$day[i] == "3"){
    data$day_num[i] = "03"
    }else if (data$day[i] == "4"){
    data$day_num[i] = "04"
    }else if (data$day[i] == "5"){
    data$day_num[i] = "05"
    }else if (data$day[i] == "6"){
    data$day_num[i] = "06"
    }else if (data$day[i] == "7"){
    data$day_num[i] = "07"
    }else if (data$day[i] == "8"){
    data$day_num[i] = "08"
    }else if (data$day[i] == "9"){
    data$day_num[i] = "09"
}else 
    data$day_num[i] = as.character(data$day[i])
}

return(data)
}


#add date variable to dataframe
add_date <- function(data){
    date_data <- make_date(data)
    data$date <- as.Date(paste(date_data$year, date_data$month_num, date_data$day_num, sep = "-" ))
    return(data)
}


```
