---
title: "Project"
author: "Group 14"
date: "April 24, 2017"
output: html_document
---

#API Coding  
We used three different APIs to add location and US Census data to our dataset.

####APIs used  
* **National Broadband Map Census API**   
    + input latitude and longitude to return corresponding census tract  
* **Google Maps API**  
    + input address to return corresponding county  
* **US Census API**   
    + input state and county to return census data  


```{r setup, warning=FALSE, message=F}
library(jsonlite)
library(stringr)
library(tidyverse)
library(readr)

data <- read_csv('guardian_data.csv')
getwd()


```


##National Broadband Map Census API 
This function uses writes a URL that requests the census tract of of a location given its latitude and longitude.  
```{r}
row_to_url <- function(row){
    # set the URL base
    basestring = "https://www.broadbandmap.gov/broadbandmap/census/tract?"
    endstring = "&format=json"
    
    x = paste0("latitude=", row$lat)
    y = paste0("&", "longitude=", row$long)
    
    # create url string from base, coordinate, and end parts        
    urlstring = paste0(basestring, x, y, endstring)
    
    return(urlstring)
}
```


Here we request some sample data. 
```{r, cache=F}
url_1 = row_to_url(data[1, ])
url_1

#We can use the `readLines( )` function to retrieve data from the API by calling `readLines(aURL)`.
dat = readLines(url_1)
dat

#The value `dat` is in string format, we use the `jsonlite` to package to convert.
datdf = fromJSON(dat)
datdf

#The Results can be accessed as follows.
datdf$Results$censusTract

#Fips
datdf$Results$censusTract$fips

#And there appears to be a useful location type field here.
#datdf$results$geometry$location_type
```


The following functions run the API requests for every datapoint at once and adds the results to our dataframe. 
```{r, warning=FALSE}
get_and_parse <- function(urlstring) {
    dat = readLines(urlstring)
    datdf = fromJSON(dat)
    return(datdf)
}

add_row <- function(df, jsndat, reqdat) {
    # expects df to have columns [street, city, state, lat, lon, loctype]
    # req dat is a row of data used for request [street, city, state]
    resdf = data.frame(
        name = reqdat$name,
        age =  reqdat$age,
        gender = reqdat$gender,               
        raceethnicity = reqdat$raceethnicity,       
        month = reqdat$month,              
        day = reqdat$day,
        year = reqdat$year,                
        streetaddress = reqdat$streetaddress,      
        city = reqdat$city,                
        state = reqdat$state,              
        classification = reqdat$classification,      
        lawenforcementagency = reqdat$lawenforcementagency,
        disposition = reqdat$disposition,          
        lat = reqdat$lat,                 
        long = reqdat$long,                
        armed = reqdat$armed, 
        state_fp = jsndat$Results$censusTract$stateFips,
        geo_id = jsndat$Results$censusTract$fips,
        census_tract = jsndat$Results$censusTract$name
    )
    return(rbind(df, resdf))
}

map_data <- function(df) {
    # expects df to have columns [street, city, state]
    # initialize a dataframe with appropriate columns using the first row of
    # our request data
    reqrow = df[1,]
    print(reqrow)
    
    qryurl = row_to_url(reqrow)
    print(qryurl)
    jsndf = get_and_parse(qryurl)
    print(jsndf)

    newdf = data.frame(
        name = reqrow$name,
        age =  reqrow$age,
        gender = reqrow$gender,               
        raceethnicity = reqrow$raceethnicity,       
        month = reqrow$month,              
        day = reqrow$day,
        year = reqrow$year,                
        streetaddress = reqrow$streetaddress,      
        city = reqrow$city,                
        state = reqrow$state,              
        classification = reqrow$classification,      
        lawenforcementagency = reqrow$lawenforcementagency,
        disposition = reqrow$disposition,          
        lat = reqrow$lat,                 
        long = reqrow$long,                
        armed = reqrow$armed, 
        state_fp = jsndf$Results$censusTract$stateFips,
        geo_id = jsndf$Results$censusTract$fips,
        census_tract = jsndf$Results$censusTract$name
    )

    # populate the dataframe with the rest of the request data
    for (ix in 2:nrow(df)) {
        tryCatch({
            reqrow = df[ix, c("name","age","gender","raceethnicity","month","day","year","streetaddress","city","state",
                          "classification","lawenforcementagency","disposition","lat","long","armed" )]

            if (!is.null(reqrow[15]) | !is.null(reqrow[16])){
                qryurl <-  row_to_url(reqrow)
                jsndf <-  get_and_parse(qryurl)
                newdf <-  add_row(newdf, jsndf, reqrow)
            }
        })
    }

    return(newdf)
}

#shortDatacomplete <- map_data(shortData)
completeData <- map_data(data)
```


##Google Maps API
This function requests a locations's county given its address and then adds it to the dataframe. 
```{r}
countyFinder <- function(row){
    
    address <- row[1]
    city <- row[2]
    state <- row[3]
    url <- paste0("https://maps.googleapis.com/maps/api/geocode/json?address=", address, ",+",
                  city, ",+", state, "&key=AIzaSyAAo44VJYaam-_y4Hs7pZic5fUZUDGk53E")
    
    raw_json <- readLines(url)
    
    # turn the json into a nice list
    info <- fromJSON(raw_json)
    
    # Get the first element of from the results list, and then the first componenent of address component
    components <- info["results"][[1]][["address_components"]][[1]]
    # Check component is not null
    if(!is.null(components)){
        # Check each elelement in address component to find the one that is listed as "administrative_area_level_2"
        # This component will give us the county name for a given request
        for (i in 1:nrow(components)){
            if (str_detect(components[i,]["types"], "administrative_area_level_2")){
                return(unlist(components[i,][["long_name"]]))
            }
        }
    }
}

importantCols <- dplyr::select(data, streetaddress, city, state)
counties <- apply(importantCols, 1, countyFinder)
counties1 <- as.character(counties)
completeData <- add_column(completeData,county = (counties1))
```



##US Census API
This followin functions request certain US Census data for each location in the dataset.  
```{r, warning=FALSE}
#Initialize data frames for important variables
medianIncomeByTract <- data.frame()


# Create base url string
    start_url <- 'http://api.census.gov/data/2015/acs5/profile?get='
    end_url <- ',NAME&for=tract:*&in=state:'
    key <- '&key=913ff64a27817f08554c0606053c0715f1de4a06'

    
# Get median Household Income    
for (i in 1:56){
      tryCatch({
        medianIncomeJSON <- readLines(paste0(start_url,'DP03_0062E',end_url,i,key))
        medianIncome <- fromJSON(medianIncomeJSON)
        medianIncome <- data.frame(medianIncome)
        medianIncome <- filter(medianIncome, X2 != 'NAME')
        medianIncome <- mutate(medianIncome, tract = unlist(lapply(str_extract_all(X2, "[0-9\\.]+"), `[[`, 1))) 
        medianIncome <- mutate(medianIncome, county1 = unlist(lapply(str_extract_all(X2, "(?<=,)[^,]+(?=,)"), `[[`, 1))) %>% select(X1, tract, county1, X3) 
        medianIncomeByTract <- rbind.data.frame(medianIncomeByTract, medianIncome)
    }, error = function(err) {
        i <- i+1
    })
}
medianIncomeByTract <-  mutate(medianIncomeByTract,county = str_trim(county1))
completeData <- left_join(completeData, medianIncomeByTract, by=c('census_tract'='tract', 'state_fp'= 'X3', "county" = "county")) %>% select(-county1)
names(completeData)[names(completeData) == 'X1'] <- 'median_household_Income'


```

We find the demographics of each census tract by looking at the breakdown of race percentages. Note, we are not using all races listed on the census, and we are also double counting the people that identify as more than one race.
```{r, warning=FALSE}
# Get percent white population  
percentWhiteByTract <- data.frame()
for (i in 1:56){
      tryCatch({
        percentWhiteJSON <- readLines(paste0(start_url,'DP05_0032PE',end_url,i,key))
        percentWhite <- fromJSON(percentWhiteJSON)
        percentWhite <- data.frame(percentWhite)
        percentWhite <- filter(percentWhite, X2 != 'NAME')
        percentWhite <- mutate(percentWhite, tract = unlist(lapply(str_extract_all(X2, "[0-9\\.]+"), `[[`, 1))) 
        percentWhite <- mutate(percentWhite, county1 = unlist(lapply(str_extract_all(X2, "(?<=,)[^,]+(?=,)"), `[[`, 1))) %>% select(X1, tract, county1, X3) 
        percentWhiteByTract <- rbind.data.frame(percentWhiteByTract,percentWhite)
    }, error = function(err) {
        i <- i+1
    })
}
percentWhiteByTract <-  mutate(percentWhiteByTract,county = str_trim(county1))
completeData <- left_join(completeData, percentWhiteByTract, by=c('census_tract'='tract', 'state_fp'= 'X3', "county" = "county")) %>% select(-county1)
names(completeData)[names(completeData) == 'X1'] <- 'percent_white_population'
```


```{r, warning=FALSE}
# Get percent black population  

percentBlackByTract <- data.frame()
for (i in 1:56){
      tryCatch({
        percentBlackJSON <- readLines(paste0(start_url,'DP05_0033PE',end_url,i,key))
        percentBlack <- fromJSON(percentBlackJSON)
        percentBlack <- data.frame(percentBlack)
        percentBlack <- filter(percentBlack, X2 != 'NAME')
        percentBlack <- mutate(percentBlack, tract = unlist(lapply(str_extract_all(X2, "[0-9\\.]+"), `[[`, 1))) 
        percentBlack <- mutate(percentBlack, county1 = unlist(lapply(str_extract_all(X2, "(?<=,)[^,]+(?=,)"), `[[`, 1))) %>% select(X1, tract, county1, X3) 
        percentBlackByTract <- rbind.data.frame(percentBlackByTract,percentBlack)
    }, error = function(err) {
        i <- i+1
    })
}
percentBlackByTract <-  mutate(percentBlackByTract,county = str_trim(county1))
completeData <- left_join(completeData, percentBlackByTract, by=c('census_tract'='tract', 'state_fp'= 'X3', "county" = "county")) %>% select(-county1)
names(completeData)[names(completeData) == 'X1'] <- 'percent_black_population'
```


```{r, warning=FALSE}
# Get percent Hispanic population  


percentHispanicByTract <- data.frame()

for (i in 1:56){
      tryCatch({
        percentHispanicJSON <- readLines(paste0(start_url,'DP05_0066PE',end_url,i,key))
        percentHispanic <- fromJSON(percentHispanicJSON)
        percentHispanic <- data.frame(percentHispanic)
        percentHispanic <- filter(percentHispanic, X2 != 'NAME')
        percentHispanic <- mutate(percentHispanic, tract = unlist(lapply(str_extract_all(X2, "[0-9\\.]+"), `[[`, 1))) 
        percentHispanic <- mutate(percentHispanic, county1 = unlist(lapply(str_extract_all(X2, "(?<=,)[^,]+(?=,)"), `[[`, 1))) %>% select(X1, tract, county1, X3) 
        percentHispanicByTract <- rbind.data.frame(percentHispanicByTract,percentHispanic)
    }, error = function(err) {
        i <- i+1
    })
}
percentHispanicByTract <-  mutate(percentHispanicByTract,county = str_trim(county1))
completeData <- left_join(completeData, percentHispanicByTract, by=c('census_tract'='tract', 'state_fp'= 'X3', "county" = "county")) %>% select(-county1)
names(completeData)[names(completeData) == 'X1'] <- 'percent_hispanic_population'
```

```{r, warning=FALSE}
percentHSGradsByTract <- data.frame()
# Get percent of people aged 25 or older that are high school graduate or higher
for (i in 1:56){
      tryCatch({
        percentHSGradsJSON <- readLines(paste0(start_url,'DP02_0066PE',end_url,i,key))
        percentHSGrads <- fromJSON(percentHSGradsJSON)
        percentHSGrads <- data.frame(percentHSGrads)
        percentHSGrads <- filter(percentHSGrads, X2 != 'NAME')
        percentHSGrads <- mutate(percentHSGrads, tract = unlist(lapply(str_extract_all(X2, "[0-9\\.]+"), `[[`, 1))) 
        percentHSGrads <- mutate(percentHSGrads, county1 = unlist(lapply(str_extract_all(X2, "(?<=,)[^,]+(?=,)"), `[[`, 1))) %>% select(X1, tract, county1, X3) 
        percentHSGradsByTract <- rbind.data.frame(percentHSGradsByTract,percentHSGrads)
    }, error = function(err) {
        i <- i+1
    })
}
percentHSGradsByTract <-  mutate(percentHSGradsByTract,county = str_trim(county1))
completeData <- left_join(completeData, percentHSGradsByTract, by=c('census_tract'='tract', 'state_fp'= 'X3', "county" = "county")) %>% select(-county1)
names(completeData)[names(completeData) == 'X1'] <- 'percent_25_or_older_HighSchool_grads'
```


Write finished data set to csv
```{r}
write_csv(completeData, 'completeData.csv')
```

