---

#BLOG POST ANALYSIS OF RECENT POLICE KILLINGS

By: Janine Mis, Jessica Honbarger, Akul Gupta

Course Website: https://idc9.github.io/stor390/

  Recently, we are constantly being informed about news regarding police killings over various causes that lead to many controversies within the American public and media. Statements such as "Black Lives Matter," "I can't breathe," or even "Racism Kills" appear on many clothing apparel public signs, and bodies of the thousands of people who protest the streets of various cities in the United States due to certain police killings of unharmful individuals. There has been a recent rise in connection between social justice and public health. The protests and media outrage all strive for equal dignity and hope for a world in which everyone can live peacefully. Being part of the overall public health, we have the knowledge, data, and capacity to promote and analyze the social determination to work towards an economic, social, racial, and community justice that helps the current situation of our public health. 
 
#Dataset Analysis
 
  Therefore, for our project, our group found data on 538's github page detailing a list of police killings as well as various pieces of background regarding the demographic and economic status of the location of incidents. We started out by analyzing the data and found that we overlooked that the data set only covered half of the year 2015. We then found a full data set from the Guardian that had information about all the killings in 2015, but lacked any background information other than location of incidents. We used those location of incidents provided by the original data set and the Google geocoding APIs to gather information about the county that the incidents occurred in. Furthermore, we also used the National Broadband Map Census API, distributed by the Federal Communications Commission, to provide us information about the corresponding census tract. Once we had the county and tract info appended to the data set, we could use the census API to pull information from the 2015 census. We used an API to get information about each state and then joined it to our original data by a join matching the county and tract. We then added information about the median household income, the race proportions, and the educational attainment for the locations of each incident to our data set. However, we found that about 50 out of 1055 observations did not have county matches in the Google API, so were initially unable to gather census data for these. Due to this lack of information, we decided to hand code in the county variable for these 50 observations in to our data. 
We further supplemented our data by adding a region variable, a binary “is_armed” variable, a date variable, and an “in_majority” variable that states whether or not the victim was in the racial majority. 
  
#Questions

  To carry out our analysis, we came up with a few questions of our own that we could explore by using the data set: 
  
  1. Does the state average graduation rate correlate to the number of police killings per state? 
  
  2. Can we predict the race of the victim based on some combination of the variables in the data set such as city income or age of victim? 
  
  3. Is the cause of death related to whether or not the deceased was armed? 
  
  4. How are victim demographics changing over time? 
  
  5. Is a certain time of the year more likely to experience police killings than others? 
  
  6. Does race and age have an effect on whether the victim was armed? 

#Plot 1 Analysis

  To further explore and analyze some of these questions, we created plots and a shiny app which illustrates a map that locates each incident with a pin mark. Clicking on a map marker in the shiny app gives you information regarding the victim about their name, age, and date of death. The user can also further filter the data points on the map in regards to certain variables contained in the data set. The application also has the capability to create a variety of bar plots in the using the data set. In regards to the plots, we wanted to analyze the distribution of the victim ages by creating a histogram of the age count for all victims. We found that the distribution was slightly skewed to the right indicating that younger individuals may be more likely to be involved and killed by police. A visualization of this histogram is the first plot below. 

#Plot 2/3 Analysis

  We also wanted to analyze the number of people killed by the police per race through a bar graph. This visualization is the second plot illustrated below. We observed that whites were killed more than any other race at a large number of 527 when compared to the second most which were blacks killed at 266. However, we then plotted this same visualization but with victims who were unarmed. For this new unarmed plot, the number of blacks killed by the police was 68 which was fairly close to the number of whites killed by the police at 91. We could relate this to the recent protests of "Black Lives Matter" in which people outraged over unarmed blacks being killed by the police. 

#Plot 4 Analysis

  In the fourth plot below, we tried to analyze if there was a relationship between state average graduation rate and number of people killed by the police per state. From this plot, we can conclude that as the state average graduation rate increases, the number of people killed by the police per state decreases. However, an outlier lies in the bottom left of the plot. If the outlier is removed, the slope of the plot would decrease and the data points would align better with the line of best fit. We can hypothesize that in an area of high education, there will be less police killings whereas in an area of low education, there will be more police killings since crime rates could possibly be high there. 
  
#Plot 5 Analysis

  Furthermore, we wanted to analyze number of police killings per region and found that states in the Southern region of the United States, have the highest number of people killed by the police at 435 compared to the Western region, which as the second highest at 367, followed by the Midwest region at 162 and Northeast region at 91, respectively. This visualization is shown as the fifth plot below. 
  
#Plot 6 Analysis

  In addition, we created a boxplot to help us understand the range of victim’s ages per death type. We can observe that the death type, struck by vehicle, had the highest average age and greatest range of age. Police killing by gunshot had majority of ages between 25 and 50 with the lowest average age among all death types. 
  
#Plot 7 Analysis

  To answer our question, of "Is a certain time of the year more likely to experience police killings than others?", we observed from our plot below (last plot), that most of the months in the year 2015 have close to the same amount of police killings except the month of December. The later months of the summer tend to have a high rate of police killings. July had the most at 122 while December had the least at only 14. 
 
#Shiny App Analysis

  Our shiny application has two main built in features. The first feature it offers to the user is an interactive map. This map shows the location of every single police killing that had occurred in 2015. Each point on the map represents a person that was killed, and clicking on a point will bring up a popup that gives information about the deceased, as well as a few background facts about the area in which the killing occurred. This map has an additional feature that gives the user the ability to filter points based on certain criteria. They can find those who were killed that were of a given race, those that has a knife, etc. The user can filter the data that appears on the map by one to three attributes, allowing them to view incidents that fit a combination of different criteria. The second feature offered by the application is the ability to build several bar plots. The user is able select a variable they would like to investigate, and the application generates a bar plot counting the number of killings for each value the variable contains. The user can then further customize the plot by breaking up the bars shown by gender. This allows them to see the gender makeup of each criteria so they may get a better feel for the distribution.

#Conclusion

Overall, our main goal of this project was to allow people to explore data about incidents in which people lost their lives at the hands of US police. We wanted to learn about who the police killed in 2015 and about the locations in which these incidents occured without making inferences or predictions. Through exploring the data, we noted how certain groups of people are killed disproportionately more than others in 2015. The plots and figures detailed below may suggest that people of certain age groups, genders, races, and areas are more likely to be killed than others. The dataset only covered one year and contained a little over 1000 data points so it would be irresponsible to make any strong assumptions or assertions about police killings. Since the topic of race and police is so sensitive, we created our shiny application to present data to the American public with the hopes that they will gather their own valuable insights.

#Appendix

```{r setup, include=FALSE}
library(jsonlite)
library(stringr)
library(tidyverse)
library(readr)


data <- read_csv("14_data_3.csv")
```


#Plot 1
```{r, echo = FALSE}

data_age <- na.omit(select(data,age))
ggplot(data_age) + geom_histogram(aes(x=age), fill = "#FF9999", color = "black", binwidth = 4) + 
    ggtitle('Police Killings by Age') + labs(x = 'Age', y = '# of people killed by police') +
    theme(panel.background = element_blank(), 
          axis.line = element_line(colour = "black", size=.1))+
    scale_x_continuous(breaks=c(0,20, 40, 60, 80),
                       limits=c(0,90),
                       expand=c(0, 1)) +
    scale_y_continuous(breaks=c(0, 50, 100, 150),
                       limits=c(0, 150),
                       expand=c(0,0)) 
```
  
  
#Plot 2
```{r, echo = FALSE}

race_count <- data %>% group_by(raceethnicity) %>% count()
ggplot(race_count) + geom_bar(aes(x=raceethnicity, y=n), stat = "identity", fill="#FF9999") + 
    theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1), axis.line = element_line(colour = "black", size=.1),
          panel.background = element_blank()) +
    ggtitle('Police Killings by Race/Ethnicity') +
    labs(x = 'Race/Ethnicity', y = '# of people killed by police') + geom_text(aes(x=raceethnicity,y=n, label=n), hjust=0.5, vjust=-.5, colour="red") 
```


#Plot 3
```{r, echo = FALSE }
race_count <- data %>% filter(is_armed==0) %>% group_by(raceethnicity) %>% count()
ggplot(race_count) + geom_bar(aes(x=raceethnicity, y=n), stat = "identity", fill="#FF9999") + 
    theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1), axis.line = element_line(colour = "black", size=.1),
          panel.background = element_blank()) +
    ggtitle('Police Killings of Unarmed People') +
    labs(x = 'Race/Ethnicity', y = '# of people killed by police') + geom_text(aes(x=raceethnicity,y=n, label=n), hjust=0.5, vjust=-.5, colour="red") 
```
  



#Plot 4
```{r ,warning=FALSE,, echo = FALSE}
avgGrad <- data %>% group_by(state) %>% summarise(avgPercentHighSchoolGrad = mean(as.numeric(percent_25_or_older_HighSchool_grads), na.rm = TRUE))
stateCount <- data %>% group_by(state) %>% summarise(DeathsPerState = n())
gradsCount <- left_join(stateCount, avgGrad, by = c('state' = 'state'))


ggplot(gradsCount) + 
    geom_point(mapping = aes(x = avgPercentHighSchoolGrad, y = DeathsPerState), color="#FF9999") +
    ggtitle('Death counts per state are higher in states with lower percentage graduation rates') +
    theme(axis.line = element_line(colour = "black", size=.1), 
          panel.background = element_blank()) +
    labs(x = 'State average graduation rate', y = '# of people killed by police per state') 
```


#Plot 5
```{r, echo = FALSE}
region_count <- data %>% group_by(region) %>% count()
ggplot(region_count) + 
    geom_bar(aes(x=region, y=n), stat = "identity", fill="#FF9999") + 
    ggtitle('Police Killings by Region') +
    labs(x = 'Region', y = '# of people killed by police') + 
    theme(axis.line = element_line(colour = "black", size=.1), 
          panel.background = element_blank()) + geom_text(aes(x=region,y=n, label=n), hjust=0.5, vjust=-.5, colour="red") 
```


#Plot 6
```{r, warning=FALSE, echo = FALSE}
Death_Type <- factor(data$classification)
ggplot(data) + geom_boxplot(aes(x=classification, y=age, fill = Death_Type)) + ggtitle('Boxplot of Age per Death Type') +
labs(x = 'Death Type', y = 'Age') + 
scale_x_discrete(limit = c("Death in custody", "Gunshot", "Struck by vehicle", "Taser"))
```

#Plot 7
```{r, echo = FALSE}
date_count <- data %>% group_by(month_num) %>% count()
ggplot(date_count) + 
    geom_bar(aes(x=month_num, y=n), stat = "identity", fill="#FF9999") + 
    ggtitle('Police Killings in 2015') +
    labs(x = 'Month', y = '# of people killed by police') + 
    theme(axis.line = element_line(colour = "black", size=.1), 
          panel.background = element_blank()) +
    scale_x_discrete(labels=c("Jan","Feb","Mar","Apr","May","Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) + geom_text(aes(x=month_num,y=n, label=n), hjust=0.5, vjust=-.5, colour="red") 
```

